\texttt{SuperScreen}: An open-source package for simulating the magnetic response of two-dimensional superconducting devices

### 2022-03-11 main.tex
overleaf changes

\author[PHYS,SIMES]{Logan Bishop-Van Horn}\ead{lbvh@stanford.edu}
\author[PHYS,AP,SIMES,GLAM]{Kathryn A. Moler\corref{kam}}\ead{kmoler@stanford.edu}

\cortext[kam]{Corresponding author.}

\address[PHYS]{Department of Physics, Stanford University, Stanford, California 94305, USA}
\address[AP]{Department of Applied Physics, Stanford University, Stanford, California 94305, USA}
\address[SIMES]{Stanford Institute for Materials and Energy Sciences, SLAC National Accelerator Laboratory, 2575 Sand Hill Road, Menlo Park, California 94025, USA}
\address[GLAM]{Geballe Laboratory for Advanced Materials, Stanford University, Stanford, California 94305, USA}

\begin{abstract}
Quantitative understanding of the spatial distribution of magnetic fields and Meissner screening currents in two-dimensional (2D) superconductors and mesoscopic thin film superconducting devices is critical to interpreting the results of magnetic measurements of such systems. Here, we introduce \SuperScreen, an open-source Python package for simulating the response of 2D superconductors to applied time-independent magnetic fields and trapped flux. Given an applied magnetic field, \SuperScreen solves the 2D London equation using an efficient matrix inversion method~\cite{Brandt2004-ew, Brandt2005-wj} to obtain the Meissner currents and magnetic fields in and around structures composed of one or more superconducting thin films of arbitrary geometry with spatially inhomogeneous magnetic penetration depth. \SuperScreen can be used to model screening effects and calculate self- and mutual-inductance in superconducting devices, and simulate the magnetic response of inhomogeneous 2D superconductors.
\end{abstract}

@@

\begin{small}
\noindent
{\em SuperScreen}\\
{\em CPC Library link to program files:} (to be added by Technical Editor) \\
{\em Developer's repository link:} \href{http://www.github.com/loganbvh/superscreen}{www.github.com/loganbvh/superscreen}\\
{\em Code Ocean capsule:} (to be added by Technical Editor)\\
{\em Licensing provisions:} MIT\\
{\em Programming language: Python}\\
% {\em Supplementary material:}\\
  % Fill in if necessary, otherwise leave out.
% {\em Journal reference of previous version:}*\\
  %Only required for a New Version summary, otherwise leave out.
% {\em Does the new version supersede the previous version?:}*\\
  %Only required for a New Version summary, otherwise leave out.
% {\em Reasons for the new version:*}\\
  %Only required for a New Version summary, otherwise leave out.
% {\em Summary of revisions:}*\\
  %Only required for a New Version summary, otherwise leave out.
{\em Nature of problem:} \SuperScreen solves for Meissner screening currents in structures composed of 2D or thin film superconductors in the presence of an applied magnetic field, pinned vortices, and trapped flux.\\
  %Describe the nature of the problem here. \\
{\em Solution method:} This package solves the 2D London equation for superconducting thin films using a matrix inversion method~\cite{Brandt2004-ew,Brandt2005-wj}.\\
  %Describe the method solution here.
% {\em Additional comments including restrictions and unusual features (approx. 50-250 words):}\\
  %Provide any additional comments here.
   \\

@@

\SuperScreen is a Python package developed to simulate the static magnetic response of structures composed of one or more layers containing superconducting thin films. \SuperScreen solves the coupled Maxwell's and London's equations in and around superconducting films with spatially-varying penetration depth in the presence of inhomogeneous applied magnetic fields, pinned vortices, and trapped flux using a matrix inversion method introduced by Brandt and Clem~\cite{Brandt2004-ew,Brandt2005-wj} and subsequently used by Kirtley, \emph{et al.} to model the magnetic response of scanning superconducting quantum interference device (SQUID) sensors.~\cite{Kirtley2016-zz, Kirtley2016-gt}

@@

It is important to note that the assumption $\vec{j}(x, y, z)\approx\vec{j}_{z_0}(x, y)$ is valid for only films that are thinner than their London penetration depth ($d<\lambda$, such that $\Lambda=\lambda^2/d>\lambda$). However the model has been applied with some success in structures with $\lambda\lesssim d$~\cite{Kirtley2016-zz,Kirtley2016-gt}. Aside from this limitation, the method described below can be used to model films with any effective penetration depth $0\leq\Lambda<\infty$.

@@

There are several important limitations to the applicability of \SuperScreen and the matrix inversion method on which it is based~\cite{Brandt2004-ew,Brandt2005-wj}. First, strictly speaking all superconducting films should be in the 2D limit, with London penetration depth $\lambda$ less than film thickness $d$, such that the current density is approximately constant along the thickness of the film. There are cases where the model reproduces experimental results despite violation of this condition (e.g. the calculations and in Refs.~\cite{Kirtley2016-zz,Kirtley2016-gt} and Section~\ref{section:examples:scanning-squid}), but care must be taking in interpreting results in these cases. Second, the model assumes that all superconducting films behave linearly and without dissipation, and that the applied magnetic field and current density are well below the critical field and critical current density of all films in a device. Third, \SuperScreen does not support ``terminal currents,'' i.e. currents flowing in one terminal of a device and out another terminal. This means that inductance calculations are limited to structures with holes, in which all applied currents are circulating currents associated with trapped flux. Terminal currents can, however, be included in stream function-based models by setting appropriate boundary conditions~\cite{Khapaev1997-kw,Khapaev2001-xq,Khapaev2001-pw,Muller2021-ci}. An extension to the model described above that treats the magnetic response of a superconducting ring interrupted by two Josephson junctions (i.e. a SQUID) with trapped vortices and terminal currents is given in Ref.~\cite{Clem2005-ye}. Finally, care should be taken to ensure that for a given model the mesh is of sufficient density that, to within the desired precision, the results of simulations do not depend on mesh size (see \ref{appendix:numerics}).

Potential improvements to \SuperScreen include: support for terminal currents as discussed above, automated determination of solution convergence for models with multiple layers, more sophisticated mesh generation (e.g. automated local mesh refinement based on device geometry or adaptive mesh refinement based on solution convergence), integration with standard integrated circuit layout software or file formats, and further optimization for memory and CPU efficiency.

@@

As mentioned in Section~\ref{section:overview:device}, the effective penetration depth $\Lambda$ of each layer can be specified as a numeric constant or as a \inline{superscreen.Parameter} that computes $\Lambda(x, y)$, the effective penetration depth as a function of position. As a (somewhat contrived) example, Code Block~\ref{code:inhomogeneous} demonstrates how to modify the model of a ring with a slit from Code Block~\ref{code:device} and Figure~\ref{fig:ring-with-slit} such that the ring's penetration depth $\Lambda(x, y)$ increases linearly as a function of the angle between the vector $\vec{r} = (x, y)$ and the $-y$ axis (along which the slit lies). The resulting current and field distributions are shown in Figure~\ref{fig:inhomogeneous}. Note that $\Lambda(x, y)$ must vary slowly on the scale of the spacing between mesh vertices in order for results to be reliable.

\section{Numerical considerations}
\label{appendix:numerics}
While the numerical method described in Section~\ref{section:implementation} is generally quite robust, it can break down for certain extreme geometries. The stream function $g(x, y)$ represents the local magnetization or density of infinitesimal current loops. The $z$-component of the magnetic field at position $\vec{r}=(x, y, z)$ from a film $F$ with stream function $g$ lying in a plane parallel to the $x-y$ plane at vertical position $z'$ is given by (see Equations~\ref{eq:field_from_kernel} and \ref{eq:kernels}):
\begin{align}
\label{eq:hz-from-kernel}
\begin{split}
    H_z(\vec{r}) &= \int_F Q_z(\vec{r},\pvec{r}')g(x', y')\,\mathrm{d}^2r',\,\text{where}\\
    Q_z(\vec{r}, \pvec{r}') &=  \frac{2(z-z')^2-\rho^2}
            {4\pi[(z-z')^2+\rho^2]^{5/2}}
\end{split}
\end{align}
and $\rho=\sqrt{(x-x')^2+(y-y')^2}$. Eq.~\ref{eq:hz-from-kernel} is exact for a continuous stream function $g$. However, the discretized version of Eq.~\ref{eq:hz-from-kernel}, in which the double integral over the film area $F$ is replaced by a sum over triangular mesh elements, is only valid if $\delta z = z-z'$, the vertical distance between the film and the point at which the field is being evaluated, is larger than the typical distance $\delta r$ between vertices in the mesh representing the film. For $z-z'=\delta z\lesssim\delta r$, the field $H_z(\vec{r})$ resembles that of a discrete set of isolated dipoles located at the mesh vertex positions, rather than that of a continuous sheet of current (see Figure~\ref{fig:xlarge-dz-dr}({\bf b})). This can lead to unphysical results when evaluating the field very close to the surface of a film (for example using \inline{Solution.field_at_position()}), or when solving models involving multi-layer structures where the vertical spacing between layers is much smaller than the lateral extent of the films, in which case the iterative calculation (Section~\ref{section:model:multilayer}) may not converge.

>>>>

\author[PHYS,SIMES]{Logan Bishop-Van Horn\corref{lbvh}}\ead{lbvh@stanford.edu}
\author[PHYS,AP,SIMES]{Kathryn A. Moler}\ead{kmoler@stanford.edu}

\cortext[lbvh]{Corresponding author.}

\address[PHYS]{Department of Physics, Stanford University, Stanford, California 94305, USA}
\address[AP]{Department of Applied Physics, Stanford University, Stanford, California 94305, USA}
\address[SIMES]{Stanford Institute for Materials and Energy Sciences, SLAC National Accelerator Laboratory, 2575 Sand Hill Road, Menlo Park, California 94025, USA}

\begin{abstract}
Quantitative understanding of the spatial distribution of magnetic fields and Meissner screening currents in two-dimensional (2D) superconductors and mesoscopic thin film superconducting devices is critical to interpreting the results of magnetic measurements of such systems. Here, we introduce \SuperScreen, an open-source Python package for simulating the response of 2D superconductors to trapped flux and applied time-independent or quasi-DC magnetic fields for any value of the effective magnetic penetration depth, $\Lambda$. Given an applied magnetic field, \SuperScreen solves the 2D London equation using an efficient matrix inversion method~\cite{Brandt2004-ew, Brandt2005-wj} to obtain the Meissner currents and magnetic fields in and around structures composed of one or more superconducting thin films of arbitrary geometry with spatially nonuniform magnetic penetration depth. \SuperScreen can be used to model screening effects and calculate self- and mutual-inductance in superconducting devices, and simulate the magnetic response of inhomogeneous 2D superconductors.
\end{abstract}

@@

\begin{small}
\noindent
{\em SuperScreen}\\
{\em CPC Library link to program files:} (to be added by Technical Editor) \\
{\em Developer's repository link:} \href{http://www.github.com/loganbvh/superscreen}{www.github.com/loganbvh/superscreen}\\
{\em Code Ocean capsule:} (to be added by Technical Editor)\\
{\em Licensing provisions:} \href{https://opensource.org/licenses/MIT}{MIT License}\\
{\em Programming language:} Python\\
% {\em Supplementary material:}\\
  % Fill in if necessary, otherwise leave out.
% {\em Journal reference of previous version:}*\\
  %Only required for a New Version summary, otherwise leave out.
% {\em Does the new version supersede the previous version?:}*\\
  %Only required for a New Version summary, otherwise leave out.
% {\em Reasons for the new version:*}\\
  %Only required for a New Version summary, otherwise leave out.
% {\em Summary of revisions:}*\\
  %Only required for a New Version summary, otherwise leave out.
{\em Nature of problem:} \SuperScreen solves for Meissner screening currents in structures composed of 2D or thin film superconductors in the presence of an applied magnetic field, pinned vortices, and trapped flux.\\
  %Describe the nature of the problem here. \\
{\em Solution method:} This package solves the 2D London equation for superconducting thin films using a matrix inversion method~\cite{Brandt2004-ew,Brandt2005-wj}.\\
  %Describe the method solution here.
% {\em Additional comments including restrictions and unusual features (approx. 50-250 words):}\\
  %Provide any additional comments here.
   \\

@@

\SuperScreen is a Python package developed to simulate the static magnetic response of structures composed of one or more layers containing superconducting thin films characterized by a London penetration depth $\lambda$ that is large compared to the film thickness $d$. \SuperScreen solves the coupled Maxwell's and London's equations in and around superconducting films with spatially-varying penetration depth in the presence of inhomogeneous applied magnetic fields, pinned vortices, and trapped flux using a matrix inversion method introduced by Brandt and Clem~\cite{Brandt2004-ew,Brandt2005-wj} and subsequently used by Kirtley, \emph{et al.} to model the magnetic response of scanning superconducting quantum interference device (SQUID) sensors~\cite{Kirtley2016-zz, Kirtley2016-gt}.

@@

It is important to note that the assumption $\vec{j}(x, y, z)\approx\vec{j}_{z_0}(x, y)$ is valid for only films that are thinner than their London penetration depth ($d\ll\lambda$, such that $\Lambda=\lambda^2/d\gg\lambda$). However the model has been applied with some success in structures with $\lambda\lesssim d$, for example by Kirtley, \emph{et al.} in modeling the magnetic response of scanning SQUID susceptometers~\cite{Kirtley2016-zz,Kirtley2016-gt}. Aside from this limitation, the method described below can be used to model films with any effective penetration depth $0\leq\Lambda<\infty$.

@@

There are several important limitations to the applicability of \SuperScreen and the matrix inversion method on which it is based~\cite{Brandt2004-ew,Brandt2005-wj}. First, strictly speaking all superconducting films should be in the 2D limit, with London penetration depth $\lambda$ large compared to the film thickness $d$, such that the current density is approximately constant along the thickness of the film. There are cases where the model reproduces experimental results despite violation of this condition (e.g. the calculations and in Refs.~\cite{Kirtley2016-zz,Kirtley2016-gt} and Section~\ref{section:examples:scanning-squid}), but care must be taking in interpreting results in these cases. Second, the model assumes that all superconducting films behave linearly and without dissipation, and that the applied magnetic field and current density are well below the critical field and critical current density of all films in a device. Third, \SuperScreen does not support ``terminal currents,'' i.e. currents flowing in one terminal of a device and out another terminal. This means that inductance calculations are limited to structures with holes, in which all applied currents are circulating currents associated with trapped flux. Terminal currents can, however, be included in stream function-based models by setting appropriate boundary conditions~\cite{Khapaev1997-kw,Khapaev2001-xq,Khapaev2001-pw,Muller2021-ci}. An extension to the model described above that treats the magnetic response of a superconducting ring interrupted by two Josephson junctions (i.e. a SQUID) with trapped vortices and terminal currents is given in Ref.~\cite{Clem2005-ye}. Finally, care should be taken to ensure that for a given model the mesh is of sufficient density that, to within the desired precision, the results of simulations do not depend on mesh size (see \ref{appendix:numerics}).

Potential improvements to \SuperScreen include: support for terminal currents as discussed above, automated determination of solution convergence for models with multiple layers, more sophisticated mesh generation (e.g. automated local mesh refinement based on device geometry or adaptive mesh refinement based on solution convergence), integration with standard integrated circuit layout software or file formats, and further numerical optimization, including GPU acceleration.

@@

As mentioned in Section~\ref{section:overview:device}, the effective penetration depth $\Lambda$ of each layer can be specified as a numeric constant or as a \inline{superscreen.Parameter} that computes $\Lambda(x, y)$, the effective penetration depth as a function of position. As a (somewhat contrived) example, Code Block~\ref{code:inhomogeneous} demonstrates how to modify the model of a ring with a slit from Code Block~\ref{code:device} and Figure~\ref{fig:ring-with-slit} such that the ring's penetration depth $\Lambda(x, y)$ increases linearly as a function of the angle between the vector $\vec{r} = (x, y)$ and the $-y$ axis (along which the slit lies). The resulting current and field distributions are shown in Figure~\ref{fig:inhomogeneous}. Note that $\Lambda(x, y)$ must vary slowly on the scale of the spacing between mesh vertices in order for results to be reliable. For example, the simulated supercurrent density near a point-like defect (i.e small compared to the triangles in the mesh) associated with a penetration depth $\Lambda_\text{defect}$ in an otherwise uniform film with penetration depth $\Lambda_0\ll\Lambda_\text{defect}$ will depend strongly on the specifics of the mesh near the defect. One can avoid this problem by increasing the density of the mesh in regions of fast-varying $\Lambda$ by adding one or more ``abstract regions" to the \inline{Device} as described in Section~\ref{section:overview:device}.

\section{Numerical considerations}
\label{appendix:numerics}
While the numerical method described in Section~\ref{section:implementation} is generally quite robust, it can break down for certain extreme geometries. The stream function $g(x, y)$ represents the local magnetization or density of infinitesimal current loops. The $z$-component of the magnetic field at position $\vec{r}=(x, y, z)$ from a film $F$ with stream function $g$ lying in a plane parallel to the $x-y$ plane at vertical position $z'$ is given by (see Equations~\ref{eq:field_from_kernel} and \ref{eq:kernels}):
\begin{align}
\label{eq:hz-from-kernel}
\begin{split}
    H_z(\vec{r}) &= \int_F Q_z(\vec{r},\pvec{r}')g(x', y')\,\mathrm{d}^2r',\,\text{where}\\
    Q_z(\vec{r}, \pvec{r}') &=  \frac{2(z-z')^2-\rho^2}
            {4\pi[(z-z')^2+\rho^2]^{5/2}}
\end{split}
\end{align}
and $\rho=\sqrt{(x-x')^2+(y-y')^2}$. Eq.~\ref{eq:hz-from-kernel} is exact for a continuous stream function $g$. However, the discretized version of Eq.~\ref{eq:hz-from-kernel}, in which the double integral over the film area $F$ is replaced by a sum over triangular mesh elements, is only valid if $\delta z = z-z'$, the vertical distance between the film and the point at which the field is being evaluated, is large compared to the typical distance $\delta r$ between vertices in the mesh representing the film. For $z-z'=\delta z\lesssim\delta r$, the field $H_z(\vec{r})$ resembles that of a discrete set of isolated dipoles located at the mesh vertex positions, rather than that of a continuous sheet of current (see Figure~\ref{fig:xlarge-dz-dr}({\bf b})). This can lead to unphysical results when evaluating the field very close to the surface of a film (for example using \inline{Solution.field_at_position()}), or when solving models involving multi-layer structures where the vertical spacing between layers is much smaller than the lateral extent of the films, in which case the iterative calculation (Section~\ref{section:model:multilayer}) may not converge.
